---
title: "Methods"
author: "Christopher Tastad"
date: \today
bibliography: bib.bib
output:
    html_document:
        toc: yes
        toc_float: yes
        theme: paper
---



# Directories

### MSI Project Folder

#### TODO

Add MSI reference for login and use

The root project directory is:

```{r eval=F}
/home/starrt2/
```

I have a personal directory under the root. At the same level, there is also a shared directory which includes some content that is under active collaboration.

```{r eval=F}
/home/starrt2/tasta005
```

Processed data can be found in the hiseq folder under data_release. This data is mirrored in the msi file structure. Each file set is named by a unique ID which follows a naming convention given by umgc.

* Look for the leading 6 digit integer as a point of reference for the data_release folder.
* The Starr_Project sub folder is given a number which is considered the MSI and sample ID.

```{r eval=F}
/home/starrt2/data_release/umgc/hiseq
```

The Analysis directory contains the source data

* illumina-basicQC # quality control data
* cellranger-`*`
    + analysis # ?
    + filtered_gene_bc_matricies # data source that has been filtered by "true cell content"
    + raw_gene_bc_matrices # matrix data that includes unfiltered reads or all read content
    + web_summary.html # interactive cell ranger summary

An example of a complete data source location is below:

```{r eval=F}
/home/starrt2/data_release/umgc/hiseq/170627_D00635_0259_ACB8R1ANXX/Starr_Project_037/Analysis/cellranger-Starr_037/filtered_gene_bc_matrices/GRCh38
```

# Seurat Tutorial

The output of the v3 PBMC data tutorial can be found [here](/home/chris/starr_lab/ocpmi/code/seurat_tutorial/pbmc3k_tutorial.html).
