---
title: "Results"
author: "Christopher Tastad"
date: "Last Edit: `r format(Sys.time(), '%Y-%m-%d')`"
bibliography: library.bib
output:
    html_document:
        toc: yes
        toc_float: yes
        theme: paper
---

## Single Cell

### Seurat

#### pt59 Full Analysis

* Tutorial: [file:///mnt/msi/tasta005/ocpmi/results/seurat/pt59/pt59_full_analysis.html](file:///mnt/msi/tasta005/ocpmi/results/seurat/pt59/pt59_full_analysis.html)
    - `~/ocpmi/results/seurat/pt59/pt59_full_analysis.html`

***

***

## wes

### Excavator2

#### excv2 Pilot redo: 1kgp GRCh38 {.tabset .tabfade}

##### Discussion

Completed *2019-08-12*

* samples: `NA10847` (control), `NA19131` (expt) via 1kgp GRCh38; `~/ocpmi/data/bam/1kgp`
* output: `~/ocpmi/results/excavator2/pilot`
* gene_ref: `hg38.fa`, `hg38_sortmerge.bed` via UCSC, `GCA_000001405.15_GRCh38.bw` via excv2
    - .fa: `~/ocpmi/data/gene_ref/ucsc/hg38`, .bw:`~/ocpmi/methods/EXCAVATOR2_Package_v1.1.2/data`

***

##### Figures

```{r, out.width = "100%", out.height = "600px", echo = F}
knitr::include_graphics("/home/chris/ocpmi/results/excavator2/pilot/Plots/NA19131/combined.pdf")
```

***

#### excv2 Pilot: pm2/15 {.tabset .tabfade}

##### Discussion

Completed *2019-08-13*

* samples: `pm2` (control), `pm15` (expt) via ocpmi; `~/ocpmi/data/bam/wes/pilot`
* output: `~/ocpmi/results/excavator2/pilot1`
* gene_ref: `hg38.fa`, `hg38_sortmerge.bed` via UCSC, `GCA_000001405.15_GRCh38.bw` via excv2
    - .fa: `~/ocpmi/data/gene_ref/ucsc/hg38`, .bw:`~/ocpmi/methods/EXCAVATOR2_Package_v1.1.2/data`

***

##### Figures

```{r, out.width = "100%", out.height = "600px", echo = F}
knitr::include_graphics("/home/chris/ocpmi/results/excavator2/pilot1/Plots/pm2/combined.pdf")
```


### VarScan2

#### vs2 Pilot: ocpmi pm2/15 {.tabset .tabfade}

##### Discussion

Completed *2019-08-08*

* samples: `pm2.bam` (control), `pm15.bam` (expt) via ocpmi; `~/ocpmi/data/bam/wes/pilot`
* output: `~/ocpmi/results/varscan2/pilot`
* gene_ref: `hg38.fa`, via UCSC; `~/ocpmi/data/gene_ref`

***

##### Figures

```{r, out.width = "100%", out.height = "600px", echo = F}
knitr::include_graphics("/home/chris/ocpmi/results/varscan2/pilot/rplot.pdf")
```

