---
title: "CNV"
output: html_notebook
---

Initialize R, set options.
Read in human chromosome data. Annotate gene chrArm locations
```{r}
# Options
myDir <- "/Users/jason/Desktop/OneDrive - University of Minnesota/BICB/Thesis/CNV/"
ensemblGenesDir <- "/Users/jason/Desktop/OneDrive - University of Minnesota/BICB/Thesis/Data/Human/ensembl_genes/"
patientDir <- paste("/Users/jason/Desktop/OneDrive - University of Minnesota/BICB/Thesis/Data/Patients/",sep="")

#patients <- c("p59","p60","p61","p62","p64","p67","p69","p71","p72","p75","p76")
patients <- "p59"

# initialize
setwd(myDir)
library(ggplot2)
library(data.table)
library(readxl)
library(tidyr)
library(reshape2)
library(RColorBrewer)
library(gplots)

# Ingest & annotate geneData from human reference data
chr1 <- read_excel(paste(ensemblGenesDir,"Chromosome 1.xls",sep = ""))
chr2 <- read_excel(paste(ensemblGenesDir,"Chromosome 2.xls",sep = ""))
chr3 <- read_excel(paste(ensemblGenesDir,"Chromosome 3.xls",sep = ""))
chr4 <- read_excel(paste(ensemblGenesDir,"Chromosome 4.xls",sep = ""))
chr5 <- read_excel(paste(ensemblGenesDir,"Chromosome 5.xls",sep = ""))
chr6 <- read_excel(paste(ensemblGenesDir,"Chromosome 6.xls",sep = ""))
chr7 <- read_excel(paste(ensemblGenesDir,"Chromosome 7.xls",sep = ""))
chr8 <- read_excel(paste(ensemblGenesDir,"Chromosome 8.xls",sep = ""))
chr9 <- read_excel(paste(ensemblGenesDir,"Chromosome 9.xls",sep = ""))
chr10 <- read_excel(paste(ensemblGenesDir,"Chromosome 10.xls",sep = ""))
chr11 <- read_excel(paste(ensemblGenesDir,"Chromosome 11.xls",sep = ""))
chr12 <- read_excel(paste(ensemblGenesDir,"Chromosome 12.xls",sep = ""))
chr13 <- read_excel(paste(ensemblGenesDir,"Chromosome 13.xls",sep = ""))
chr14 <- read_excel(paste(ensemblGenesDir,"Chromosome 14.xls",sep = ""))
chr15 <- read_excel(paste(ensemblGenesDir,"Chromosome 15.xls",sep = ""))
chr16 <- read_excel(paste(ensemblGenesDir,"Chromosome 16.xls",sep = ""))
chr17 <- read_excel(paste(ensemblGenesDir,"Chromosome 17.xls",sep = ""))
chr18 <- read_excel(paste(ensemblGenesDir,"Chromosome 18.xls",sep = ""))
chr19 <- read_excel(paste(ensemblGenesDir,"Chromosome 19.xls",sep = ""))
chr20 <- read_excel(paste(ensemblGenesDir,"Chromosome 20.xls",sep = ""))
chr21 <- read_excel(paste(ensemblGenesDir,"Chromosome 21.xls",sep = ""))
chr22 <- read_excel(paste(ensemblGenesDir,"Chromosome 22.xls",sep = ""))
chrX <- read_excel(paste(ensemblGenesDir,"Chromosome X.xls",sep = ""))

geneData <- rbind(chr1, chr2, chr3, chr4, chr5, chr6, chr7, chr8, chr9, chr10, chr11, chr12, chr13, chr14, chr15, chr16, chr17, chr18, chr19, chr20, chr21, chr22, chrX)

colnames(geneData) <- c("ENSEMBL", "chr", "start", "end", "geneName")

rm(chr1, chr2, chr3, chr4, chr5, chr6, chr7, chr8, chr9, chr10, chr11, chr12, chr13, chr14, chr15, chr16, chr17, chr18, chr19, chr20, chr21, chr22, chrX)

geneData$arm <- "NA"
geneData$arm[geneData$chr == 1 & geneData$start < 125000000] <- 'p'
geneData$arm[geneData$chr == 1 & geneData$start > 125000000] <- 'q'
geneData$arm[geneData$chr == 2 & geneData$start < 93300000] <- 'p'
geneData$arm[geneData$chr == 2 & geneData$start > 93300000] <- 'q'
geneData$arm[geneData$chr == 3 & geneData$start < 91000000] <- 'p'
geneData$arm[geneData$chr == 3 & geneData$start > 91000000] <- 'q'
geneData$arm[geneData$chr == 4 & geneData$start < 50400000] <- 'p'
geneData$arm[geneData$chr == 4 & geneData$start > 50400000] <- 'q'
geneData$arm[geneData$chr == 5 & geneData$start < 48400000] <- 'p'
geneData$arm[geneData$chr == 5 & geneData$start > 48400000] <- 'q'
geneData$arm[geneData$chr == 6 & geneData$start < 61000000] <- 'p'
geneData$arm[geneData$chr == 6 & geneData$start > 61000000] <- 'q'
geneData$arm[geneData$chr == 7 & geneData$start < 59900000] <- 'p'
geneData$arm[geneData$chr == 7 & geneData$start > 59900000] <- 'q'
geneData$arm[geneData$chr == 8 & geneData$start < 45600000] <- 'p'
geneData$arm[geneData$chr == 8 & geneData$start > 45600000] <- 'q'
geneData$arm[geneData$chr == 9 & geneData$start < 49000000] <- 'p'
geneData$arm[geneData$chr == 9 & geneData$start > 49000000] <- 'q'
geneData$arm[geneData$chr == 10 & geneData$start < 40200000] <- 'p'
geneData$arm[geneData$chr == 10 & geneData$start > 40200000] <- 'q'
geneData$arm[geneData$chr == 11 & geneData$start < 53700000] <- 'p'
geneData$arm[geneData$chr == 11 & geneData$start > 53700000] <- 'q'
geneData$arm[geneData$chr == 12 & geneData$start < 35800000] <- 'p'
geneData$arm[geneData$chr == 12 & geneData$start > 35800000] <- 'q'
geneData$arm[geneData$chr == 13 & geneData$start < 17900000] <- 'p'
geneData$arm[geneData$chr == 13 & geneData$start > 17900000] <- 'q'
geneData$arm[geneData$chr == 14 & geneData$start < 17600000] <- 'p'
geneData$arm[geneData$chr == 14 & geneData$start > 17600000] <- 'q'
geneData$arm[geneData$chr == 15 & geneData$start < 19000000] <- 'p'
geneData$arm[geneData$chr == 15 & geneData$start > 19000000] <- 'q'
geneData$arm[geneData$chr == 16 & geneData$start < 36600000] <- 'p'
geneData$arm[geneData$chr == 16 & geneData$start > 36600000] <- 'q'
geneData$arm[geneData$chr == 17 & geneData$start < 24000000] <- 'p'
geneData$arm[geneData$chr == 17 & geneData$start > 24000000] <- 'q'
geneData$arm[geneData$chr == 18 & geneData$start < 17200000] <- 'p'
geneData$arm[geneData$chr == 18 & geneData$start > 17200000] <- 'q'
geneData$arm[geneData$chr == 19 & geneData$start < 26500000] <- 'p'
geneData$arm[geneData$chr == 19 & geneData$start > 26500000] <- 'q'
geneData$arm[geneData$chr == 20 & geneData$start < 27500000] <- 'p'
geneData$arm[geneData$chr == 20 & geneData$start > 27500000] <- 'q'
geneData$arm[geneData$chr == 21 & geneData$start < 13200000] <- 'p'
geneData$arm[geneData$chr == 21 & geneData$start > 13200000] <- 'q'
geneData$arm[geneData$chr == 22 & geneData$start < 14700000] <- 'p'
geneData$arm[geneData$chr == 22 & geneData$start > 14700000] <- 'q'
geneData$arm[geneData$chr == "X" & geneData$start < 60600000] <- 'p'
geneData$arm[geneData$chr == "X" & geneData$start > 60600000] <- 'q'

```



Read in patient data. 
Read in cluster annotation data.
Create master matrix with all cells all genes
```{r}
#for( i in 1:length(patients) )
#{
  i <- 1
  patient <- patients[i]  

  # read in files and re-label headers
  barcod
  
  es.tsv <- read.csv(paste(patientDir,patient,"/barcodes.tsv", sep=""), header = FALSE)
  colnames(barcodes.tsv) <- "barcodes"

  genes.tsv <- read.csv(paste(patientDir,patient,"/genes.tsv", sep=""), header = FALSE, sep = "\t")
  colnames(genes.tsv) <- c("ENSEMBL","Names")

  matrix.mtx <- read.csv(paste(patientDir,patient,"/matrix.mtx", sep=""), header = FALSE, skip = 3, sep = " ")
  colnames(matrix.mtx) <- c("geneIndex", "cellIndex", "UMI")
  
  if ( file.exists(paste(patientDir,patient,"/cluster_data.csv", sep="")) ) 
  { 
    clusterDataExists <- TRUE 
  } else { 
    clusterDataExists <- FALSE 
  }
  
  if ( clusterDataExists )
  { 
    cluster_data.csv <- read.csv(paste(patientDir,patient,"/cluster_data.csv", sep=""), header = TRUE, skip = 0, sep = ",") 
    colnames(cluster_data.csv) <- c("origIndex", "filterIndex", "barcode", 
                                  "clustNum", "clustColor", 
                                  "majorCellType", "majorCellTypeColor", 
                                  "minorCellType", "minorCellTypeColor", 
                                  "tSNE1", "tSNE2")
  } else { 
    annot1 <- read.csv(paste(patientDir,patient,"/annot-01.csv", sep=""), header = TRUE, sep = ",")
    annot2 <- read.csv(paste(patientDir,patient,"/annot-02.csv", sep=""), header = TRUE, sep = ",") 
    colnames(annot1) <- c("barcode", "clustNum")
    colnames(annot2) <- c("barcode", "cellType")
  }

  # add ensemble index from genes.tsv file
  geneData$ensemblIndex <- match(geneData$ENSEMBL, genes.tsv$ENSEMBL)

  # use indices of genes to match to UMI values in matrix
  masterMatrix <- merge(geneData, matrix.mtx, by.x="ensemblIndex", by.y="geneIndex", all.x=FALSE, all.y = FALSE)

  # add barcodes, cluster numbers, and cell types to masterMatrix
  masterMatrix$barcode <- barcodes.tsv[masterMatrix$cellIndex,]
  if ( clusterDataExists )
  {
    masterMatrix$clustNum <- cluster_data.csv$clustNum[match(masterMatrix$barcode, cluster_data.csv$barcode)]
    masterMatrix$cellType <- cluster_data.csv$minorCellType[match(masterMatrix$barcode, cluster_data.csv$barcode)]
  } else {
    masterMatrix$clustNum <- annot1$clustNum[match(masterMatrix$barcode, annot1$barcode)]
    masterMatrix$cellType <- annot2$cellType[match(masterMatrix$barcode, annot2$barcode)]
  }
  masterMatrix$chrArm <- paste(masterMatrix$chr, masterMatrix$arm, sep = "")
```


Filter data?
```{r}

```

Normalize data 1:
  For every UMI, find fraction of total cell's UMI.
  Sum these values per chrArm. Now we have total fraction of UMI for each cell for each chrArm
  Normalize by chrArm using only specific cell types as factors for centering and scaling
```{r}
  
  ### Turn UMI counts into fractions of total cell UMI by dividing each UMI by sum of all UMIs for that cell
  totCellUMI <- aggregate( UMI ~ barcode, data = masterMatrix, FUN = sum )
  colnames(totCellUMI) <- c("barcode", "sumUMI")
  masterMatrix <- merge(masterMatrix, totCellUMI, by.x="barcode", by.y="barcode", all.x=FALSE)
  rm(totCellUMI)
  masterMatrix$UMIfrac <- masterMatrix$UMI / masterMatrix$sumUMI

  # sum value per chrArm  
  chrArmFracSums <- aggregate( UMIfrac ~ barcode + chrArm, data = masterMatrix, FUN = sum  )
  # add cell type data
  chrArmFracSums <- merge(x = chrArmFracSums, y = cluster_data.csv, by = "barcode", all.x = FALSE, all.y=FALSE)
   
  # calculate normalized values using fibroblasts and macrophages
  centerNonEpi <- aggregate( formula = UMIfrac ~ chrArm, data = chrArmFracSums[chrArmFracSums$minorCellType == "epithelial",], FUN = mean )
  scaleNonEpi <- aggregate( formula = UMIfrac ~ chrArm, data = chrArmFracSums[chrArmFracSums$minorCellType == "epithelial",], FUN = sd )
  colnames(centerNonEpi) <- c("chrArm", "meanUMInonEpi")
  colnames(scaleNonEpi) <- c("chrArm", "sdUMInonEpi")
  
  #centerFibro <- aggregate( formula = UMIfrac ~ chrArm, data = chrArmFracSums[chrArmFracSums$minorCellType == "fibroblast",], FUN = mean )
  #centerMacro <- aggregate( formula = UMIfrac ~ chrArm, data = chrArmFracSums[chrArmFracSums$minorCellType == "macrophage",], FUN = mean )
  #scaleFibro <- aggregate( formula = UMIfrac ~ chrArm, data = chrArmFracSums[chrArmFracSums$minorCellType == "fibroblast",], FUN = sd )
  #scaleMacro <- aggregate( formula = UMIfrac ~ chrArm, data = chrArmFracSums[chrArmFracSums$minorCellType == "macrophage",], FUN = sd )
  #colnames(centerFibro) <- c("chrArm", "meanUMIfibro")
  #colnames(scaleFibro) <- c("chrArm", "sdUMIfibro")
  #colnames(centerMacro) <- c("chrArm", "meanUMImacro")
  #colnames(scaleMacro) <- c("chrArm", "sdUMImacro")
  
  chrArmFracSums <- merge(x = chrArmFracSums, y = centerFibro, by = "chrArm", all.x = FALSE, all.y = FALSE)
  chrArmFracSums <- merge(x = chrArmFracSums, y = centerMacro, by = "chrArm", all.x = FALSE, all.y = FALSE)
  chrArmFracSums <- merge(x = chrArmFracSums, y = scaleFibro, by = "chrArm", all.x = FALSE, all.y = FALSE)
  chrArmFracSums <- merge(x = chrArmFracSums, y = scaleMacro, by = "chrArm", all.x = FALSE, all.y = FALSE)
  
  chrArmFracSums$UMInormFibro <- ( chrArmFracSums$UMIfrac - chrArmFracSums$meanUMIfibro ) / chrArmFracSums$sdUMIfibro
  chrArmFracSums$UMInormMacro <- ( chrArmFracSums$UMIfrac - chrArmFracSums$meanUMImacro ) / chrArmFracSums$sdUMImacro

```

Create Fibro heatmap
```{r}

  info <- paste(patient, "total fraction of UMI counts per cell, centered/scaled by chrArm on fibroblast cells", sep = "; ")

  # convert data to wide matrix
  heatmapMatrix <- chrArmFracSums[,c("chrArm", "barcode", "UMInormFibro")] 
  heatmapMatrix <- dcast(heatmapMatrix, barcode~chrArm, value.var="UMInormFibro") 
  
  # add cell type, clust num, and color data
  heatmapMatrix$cellType <- cluster_data.csv$minorCellType[match(heatmapMatrix$barcode, cluster_data.csv$barcode, nomatch = "NA")]
  heatmapMatrix$clustNum <- cluster_data.csv$clustNum[match(heatmapMatrix$barcode, cluster_data.csv$barcode, nomatch = "NA")]
  heatmapMatrix$color    <- cluster_data.csv$minorCellTypeColor[match(heatmapMatrix$barcode, cluster_data.csv$barcode, nomatch = "NA")]

  # fix rownames
  rownames(heatmapMatrix) <- heatmapMatrix[,"barcode"] 
  
  # order y-axis (cells) by cellType, then clustNum
  heatmapMatrix <- heatmapMatrix[order(heatmapMatrix$cellType, heatmapMatrix$clustNum),]
  
  # create row seperators based on cluster numbers
  rowSeps <- c(1,1+which(diff(heatmapMatrix$clustNum)!=0))

  # save heatmap colors
  heatmapColors <- heatmapMatrix$color
  heatmapColors[heatmapColors==""] <- NA
  
  # make group row labels
  groupLabels <- rep("", nrow(heatmapMatrix))
  groupLabels[rowSeps] <- paste(heatmapMatrix$clustNum[rowSeps], heatmapMatrix$cellType[rowSeps], sep = ",")
  
  # order x-axis by chrArm, drop: "barcode","cellType","color"
  heatmapMatrix <- heatmapMatrix[,c("1p","1q","2p","2q","3p","3q","4p","4q","5p","5q","6p","6q","7p","7q","8p","8q","9p","9q","10p","10q","11p","11q","12p","12q","13q","14q","15q","16p","16q","17p","17q","18p","18q","19p","19q","20p","20q","21p","21q","22q","Xp","Xq")]
  
  heatmapMatrix<- as.matrix(x = heatmapMatrix)
  
  scaling <- "none"
  
  # create heatmap
  png(filename = paste(myDir,"heatmaps/",patient,".heatmap.fibro.png", sep=""), width = 3000, height = 1500)
  heatmap.2(x = heatmapMatrix,
            Rowv = FALSE, Colv = FALSE,
            dendrogram = "none", 
            scale = scaling,
            trace = "none",
            labRow = groupLabels,
            RowSideColors = as.character(heatmapColors),
            rowsep = rowSeps,
            main = info,
            xlab = "Chromosome Arm", ylab = "Cells",
            col = colorRampPalette(c("red", "white", "green"))(n = 299), 
            breaks = c(seq(-3,-1.5,length=100),seq(-1.49,1.49,length=100),seq(1.5,3,length=100)), 
#           breaks = c(seq(-2,-1.25,length=100),seq(-1.24,1.24,length=100),seq(1.25,2,length=100)),
            symbreaks = TRUE,
            symm = FALSE, 
            symkey = TRUE,
            keysize = 0.5,
            cexRow = 1
            )
  dev.off()

    
```



Create Macro heatmap
```{r}

  info <- paste(patient, "total fraction of UMI counts per cell, centered/scaled by chrArm on macrophage cells", sep = "; ")

  # convert data to wide matrix
  heatmapMatrix <- chrArmFracSums[,c("chrArm", "barcode", "UMInormMacro")] 
  heatmapMatrix <- dcast(heatmapMatrix, barcode~chrArm, value.var="UMInormMacro") 
  
  # add cell type, clust num, and color data
  heatmapMatrix$cellType <- cluster_data.csv$minorCellType[match(heatmapMatrix$barcode, cluster_data.csv$barcode, nomatch = "NA")]
  heatmapMatrix$clustNum <- cluster_data.csv$clustNum[match(heatmapMatrix$barcode, cluster_data.csv$barcode, nomatch = "NA")]
  heatmapMatrix$color    <- cluster_data.csv$minorCellTypeColor[match(heatmapMatrix$barcode, cluster_data.csv$barcode, nomatch = "NA")]

  # fix rownames
  rownames(heatmapMatrix) <- heatmapMatrix[,"barcode"] 
  
  # order y-axis (cells) by cellType, then clustNum
  heatmapMatrix <- heatmapMatrix[order(heatmapMatrix$cellType, heatmapMatrix$clustNum),]
  
  # create row seperators based on cluster numbers
  rowSeps <- c(1,1+which(diff(heatmapMatrix$clustNum)!=0))

  # save heatmap colors
  heatmapColors <- heatmapMatrix$color
  heatmapColors[heatmapColors==""] <- NA
  
  # make group row labels
  groupLabels <- rep("", nrow(heatmapMatrix))
  groupLabels[rowSeps] <- paste(heatmapMatrix$clustNum[rowSeps], heatmapMatrix$cellType[rowSeps], sep = ",")
  
  # order x-axis by chrArm, drop: "barcode","cellType","color"
  heatmapMatrix <- heatmapMatrix[,c("1p","1q","2p","2q","3p","3q","4p","4q","5p","5q","6p","6q","7p","7q","8p","8q","9p","9q","10p","10q","11p","11q","12p","12q","13q","14q","15q","16p","16q","17p","17q","18p","18q","19p","19q","20p","20q","21p","21q","22q","Xp","Xq")]
  
  heatmapMatrix<- as.matrix(x = heatmapMatrix)
  
  scaling <- "none"
  
  # create heatmap
  png(filename = paste(myDir,"heatmaps/",patient,".heatmap.macro.new.png", sep=""), width = 3000, height = 1500)
  heatmap.2(x = heatmapMatrix,
            Rowv = FALSE, Colv = FALSE,
            dendrogram = "none", 
            scale = scaling,
            trace = "none",
            labRow = groupLabels,
            RowSideColors = as.character(heatmapColors),
            rowsep = rowSeps,
            main = info,
            xlab = "Chromosome Arm", ylab = "Cells",
            col = colorRampPalette(c("red", "white", "blue"))(99), 
       #     breaks = c(seq(-2,-1.501,length=100),        # red-white
        #               seq(-1.5,-1.001,length=100),      # white-white
        #                seq( -1, 1, length=100),          # white-blue
        #                seq(01.001,1.5,length=100),       # white-blue
          #            ), 
            symbreaks = TRUE,
            symm = FALSE, 
            symkey = TRUE,
            keysize = 0.5,
            cexRow = 1
            )
  dev.off()

    
```


























Old code...
```{r}


  # scale and log transform fractions
  # masterMatrix$UMIscaledFrac <- log(masterMatrix$UMIfrac * 10000)
  
  # center and scale columns of data based on macro/fibro subsets
  # find average for centering for each chrArm, using only the macro/fibro cells
  centerFibro <- aggregate( formula = UMI ~ chrArm, data = masterMatrix[masterMatrix$cellType == "fibroblast",], FUN = mean )
  colnames(centerFibro) <- c("chrArm", "meanUMIfibro")
  scaleFibro <- aggregate( formula = UMI ~ chrArm, data = masterMatrix[masterMatrix$cellType == "fibroblast",], FUN = sd )
  colnames(scaleFibro) <- c("chrArm", "sdUMIfibro")
  centerMacro <- aggregate( formula = UMI ~ chrArm, data = masterMatrix[masterMatrix$cellType == "macrophage",], FUN = mean )
  colnames(centerMacro) <- c("chrArm", "meanUMImacro")
  scaleMacro <- aggregate( formula = UMI ~ chrArm, data = masterMatrix[masterMatrix$cellType == "macrophage",], FUN = sd )
  colnames(scaleMacro) <- c("chrArm", "sdUMImacro")
  
   # masterMatrix$UMInormFibro <- ( masterMatrix$UMI - centerFibro$meanUMI[centerFibro$chrArm == masterMatrix$chrArm] ) / scaleFibro$sdUMI[scaleFibro$chrArm == masterMatrix$chrArm]  ## WHY DOESN'T THIS WORK? Adding a ton of repetitive data points to masterMatrix to get around this : ( 
  
  masterMatrix <- merge(x = masterMatrix, y = centerFibro, by = "chrArm", all.x = FALSE)
  masterMatrix <- merge(x = masterMatrix, y = scaleFibro, by = "chrArm", all.x = FALSE)
  masterMatrix <- merge(x = masterMatrix, y = centerMacro, by = "chrArm", all.x = FALSE)
  masterMatrix <- merge(x = masterMatrix, y = scaleMacro, by = "chrArm", all.x = FALSE)

  masterMatrix$UMInormFibr <- ( masterMatrix$UMI - masterMatrix$meanUMIfibro ) / masterMatrix$sdUMIfibro
  masterMatrix$UMInormMacro <- ( masterMatrix$UMI - masterMatrix$meanUMImacro ) / masterMatrix$sdUMImacro

  ###
  ### create matrix for heatmap
  ###
  
    ###  get the average UMI for all genes in each cell / chr arm.
    # heatmapData <- aggregate( UMI~barcode+chrArm, data = masterMatrix, FUN = mean ) 
    # info <- paste(patient, "mean of UMI counts", sep = "; ")
    ## OR ## 
    # get the sum of UMI fractions, where UMI fractions are total UMI count of gene / total UMI count of cell
    # heatmapData <- aggregate( UMIfrac~barcode+chrArm, data = masterMatrix, FUN = sum ) 
    # info <- paste(patient, "sum of UMI fractions", sep = "; ")
    ## OR ##
    # get the normalized values
    # heatmapData <- aggregate( UMInormMacro~barcode+chrArm, data = masterMatrix, FUN = mean ) 
    # info <- paste(patient, "mean of UMI values centered/scaled by chrArm on macrophage cells", sep = "; ")
    ## OR ##
    heatmapData <- aggregate( UMInormFibr ~ barcode+chrArm, data = masterMatrix, FUN = mean ) 
    info <- paste(patient, "mean of UMI values centered/scaled by chrArm on fibroblast cells", sep = "; ")
    
     
     
     
  # add cell type 
  if ( clusterDataExists )
  {
    heatmapData$cellType <- cluster_data.csv$minorCellType[match(heatmapData$barcode, cluster_data.csv$barcode)]
  } else {    
    heatmapData$cellType <- annot2$cellType[match(heatmapData$barcode, annot2$barcode)] 
  }
  
  # pull out data for matrix
  heatmapMatrix <- heatmapData[,1:3] 
  
  # convert to wide matrix
  # heatmapMatrix <- dcast(heatmapMatrix, barcode~chrArm, value.var="UMI") 
  # heatmapMatrix <- dcast(heatmapMatrix, barcode~chrArm, value.var="UMIfrac") 
  # heatmapMatrix <- dcast(heatmapMatrix, barcode~chrArm, value.var="UMInormMacro") 
  heatmapMatrix <- dcast(heatmapMatrix, barcode~chrArm, value.var="UMInormFibr") 
  
  
  # add cell type & clust num data
  if ( clusterDataExists )
  {  
    heatmapMatrix$cellType <- cluster_data.csv$minorCellType[match(heatmapMatrix$barcode, cluster_data.csv$barcode)]
    heatmapMatrix$clustNum <- cluster_data.csv$clustNum[match(heatmapMatrix$barcode, cluster_data.csv$barcode)]
  } else {
    heatmapMatrix$cellType <- annot2$cellType[match(heatmapMatrix$barcode, annot2$barcode)]
    heatmapMatrix$clustNum <- annot1$clustNum[match(heatmapMatrix$barcode, annot1$barcode)]
  }
  
  # fix row names
  rownames(heatmapMatrix) <- heatmapMatrix[,"barcode"] 
  
  # order y-axis (cells) by cellType, then clustNum
  heatmapMatrix <- heatmapMatrix[order(heatmapMatrix$cellType, heatmapMatrix$clustNum),]
  
  # create row seperators based on cluster numbers
  rowSeps <- c(1,1+which(diff(heatmapMatrix$clustNum)!=0))
  #colSeps <- seq(from = 1, to = 44, by = 2)
  
  # add colors
  if ( clusterDataExists )
  {  
    heatmapMatrix$color <- cluster_data.csv$minorCellTypeColor[match(heatmapMatrix$barcode, 
                                                                     cluster_data.csv$barcode, 
                                                                     nomatch = "NA")]
  } else {
    heatmapMatrix$color[heatmapMatrix$cellType == "epithelial"] <- "red" 
    heatmapMatrix$color[heatmapMatrix$cellType == "immune"] <- "yellow"
    heatmapMatrix$color[heatmapMatrix$cellType == "stroma"] <- "blue" 
    heatmapMatrix$color[heatmapMatrix$cellType == "endothelial"] <- "green" 
  }
  heatmapColors <- heatmapMatrix$color
  heatmapColors[heatmapColors==""] <- NA
  
  # make group row labels
  groupLabels <- rep("", nrow(heatmapMatrix))
  groupLabels[rowSeps] <- paste(heatmapMatrix$clustNum[rowSeps], heatmapMatrix$cellType[rowSeps], sep = ",")
  
  # order x-axis by chrArm, drop: "barcode","cellType","color"
  heatmapMatrix <- heatmapMatrix[,c("1p","1q","2p","2q","3p","3q","4p","4q","5p","5q","6p","6q","7p","7q","8p","8q","9p","9q","10p","10q","11p","11q","12p","12q","13q","14q","15q","16p","16q","17p","17q","18p","18q","19p","19q","20p","20q","21p","21q","22q","Xp","Xq")]
  
  # round to 2 digits
  # heatmapMatrix<- as.matrix(round(x = heatmapMatrix, digits = 2))
  heatmapMatrix<- as.matrix(x = heatmapMatrix)
  
  scaling <- "none"
  info <- paste(info, scaling, sep = "; ")
  
  # create heatmap
  png(filename = paste(myDir,"heatmaps/",patient,".heatmap.4.png", sep=""), width = 3000, height = 1500)
  heatmap.2(x = heatmapMatrix,
            Rowv = FALSE, Colv = FALSE,
            dendrogram = "none", 
            scale = scaling,
            trace = "none",
            labRow = groupLabels,
            RowSideColors = as.character(heatmapColors),
            rowsep = rowSeps,
            main = info,
            xlab = "Chromosome Arm", ylab = "Cells",
            col = colorRampPalette(c("red", "black", "green"))(n = 299), 
            breaks = c(seq(-3,-1,length=100),seq(-0.99,0.99,length=100),seq(1,3,length=100)), 
#           breaks = c(seq(-2,-1.25,length=100),seq(-1.24,1.24,length=100),seq(1.25,2,length=100)),
            symbreaks = TRUE,
            symm = FALSE, 
            symkey = TRUE,
            keysize = 0.5,
            cexRow = 1
            )
  dev.off()
```
}








How does our genes.tsv gene list compare with BioMart gene list?
```{r}
nrow(genes.tsv)
nrow(geneData)
length(union(genes.tsv$ENSEMBL, geneData$ENSEMBL))
length(intersect(genes.tsv$ENSEMBL, geneData$ENSEMBL))
length(setdiff(genes.tsv$ENSEMBL, geneData$ENSEMBL))
length(setdiff(geneData$ENSEMBL, genes.tsv$ENSEMBL))
```



Now into plotting...

```{r}
# How many cells is each gene expressed in?
DT <- data.table(masterMatrix)

geneData1 <- DT[, .(numCells = uniqueN(cellIndex)), by = ENSEMBL]
geneData2 <- DT[, .(UMIAve = mean(UMI)), by = ENSEMBL]
geneData <- merge(geneData1, geneData2)
chr8matrix <- merge(chr8matrix, geneData)

# How many genes does each cell have expression for?
cellData1 <- DT[, .(numGenes = uniqueN(ENSEMBL)), by = cellIndex]
cellData2 <- DT[, .(cellType = cellType), by = cellIndex]
cellData <- unique(merge(cellData1, cellData2))
chr8matrix <- merge(chr8matrix, cellData)
rm(cellData1, cellData2)

pArm <- chr8matrix[chr8matrix$pArm == TRUE,]
pArmDT <- data.table(pArm)
pArmGeneData1 <- pArmDT[, .(numCells = uniqueN(cellIndex)), by = ENSEMBL]
pArmGeneData2 <- pArmDT[, .(UMIAve = mean(UMI)), by = ENSEMBL]
pArmgeneData <- merge(pArmGeneData1, pArmGeneData2)
pArmCellData1 <- pArmDT[, .(numGenes = uniqueN(ENSEMBL)), by = cellIndex]
pArmCellData2 <- pArmDT[, .(cellType = cellType), by = cellIndex]
pArmCellData <- unique(merge(pArmCellData1, pArmCellData2))

qArm <- chr8matrix[chr8matrix$pArm == FALSE,]
qArmDT <- data.table(qArm)
qArmGeneData1 <- qArmDT[, .(numCells = uniqueN(cellIndex)), by = ENSEMBL]
qArmGeneData2 <- qArmDT[, .(UMIAve = mean(UMI)), by = ENSEMBL]
qArmgeneData <- merge(qArmGeneData1, qArmGeneData2)
qArmCellData1 <- qArmDT[, .(numGenes = uniqueN(ENSEMBL)), by = cellIndex]
qArmCellData2 <- qArmDT[, .(cellType = cellType), by = cellIndex]
qArmCellData <- unique(merge(qArmCellData1, qArmCellData2))

rm(pArm, pArmDT)
rm(geneData1, geneData2, pArmGeneData1, pArmGeneData2, qArmGeneData1, qArmCellData2)

```


In how many cells is each gene expressed?
```{r}
ggplot(data = geneData) + 
  geom_histogram(mapping = aes(x=numCells), binwidth = 50) +
  ggtitle("Num cells per gene") + 
  xlab("num cells") +
  ylab("num genes")
```


What is the average UMI for each gene?
```{r}
ggplot(data = geneData) + 
  geom_histogram(mapping = aes(x=UMIAve), binwidth = 1) + 
  ggtitle("Ave(UMI) per gene") +
  xlab("ave(UMI)") +
  ylab("num genes")
```

Find out how many genes each cell has expression in
```{r}
ggplot(data = cellData,
       aes(x = numGenes,
           fill = cellType)) + 
  geom_histogram(binwidth = 5,
                 position = "dodge") +
  ggtitle("Total num genes per cell") 
  xlab("num genes") +
  ylab("num cells")
```

How many genes per cell on the p arm?
```{r}
ggplot(data = pArmCellData,
       aes(x = numGenes,
          fill = cellType)) +
  geom_histogram(position = "dodge", binwidth = 5) +
  ggtitle("num genes per cell - p arm") +
  xlab("num genes") +
  ylab("num cells")
```

How many genes per cell on the q arm? 
```{r}
ggplot(data = qArmCellData,
       aes(x = numGenes,
          fill = cellType)) +
  geom_histogram(position = "dodge", binwidth = 5) +
  ggtitle("num genes per cell - q arm") +
  xlab("num genes") +
  ylab("num cells")
```

Is the distribution of UMI occurrence bimodal in chr8p?
```{r}
# ggplot(data = pArmNumGenesByCell) + 
#   geom_histogram(mapping = aes(x=numGenes), binwidth = 5) + 
#   ggtitle("num genes per cell - p arm") +
#   xlab("num genes") +
#   ylab("num cells")

```

Is the distribution of UMI occurrence bimodal in chr8q? No
```{r}
# qArm <- chr8matrix[chr8matrix$pArm == FALSE,]
# qArmDT <- data.table(qArm)
# qArmNumGenesByCell <- qArmDT[, .(numGenes = uniqueN(ENSEMBL)), by = cellIndex]
# nrow(qArmNumGenesByCell)
# ggplot(data = qArmNumGenesByCell) + 
#   geom_histogram(mapping = aes(x=numGenes), binwidth = 5) + 
#   ggtitle("num genes per cell - q arm") +
#   xlab("num genes") +
#   ylab("num cells")

```

What is the distribution of UMI sums for cells for chr8p only?
```{r}
# sumTableP <- aggregate(UMI~cellIndex, data=pArm, FUN=sum)
# colnames(sumTableP) <- c("cellIndex", "sumUMI")
# max(sumTableP$sumUMI)
# min(sumTableP$sumUMI)
# ggplot(data = sumTableP) + 
#   geom_histogram(mapping = aes(x=sumUMI), binwidth = 5) + 
#   ggtitle("Sum(UMI) per cell - p arm") + 
#   xlab("sum(UMI) - p arm only") +
#   ylab("num cells")

```

What is the distribution of UMI sums for cells for chr8q only?
```{r}
# sumTableQ <- aggregate(UMI~cellIndex, data=qArm, FUN=sum)
# colnames(sumTableQ) <- c("cellIndex", "sumUMI")
# max(sumTableQ$sumUMI)
# min(sumTableQ$sumUMI)
# ggplot(data = sumTableQ) + 
#   geom_histogram(mapping = aes(x=sumUMI), binwidth = 5) + 
#   ggtitle("Distribution of sum(UMI) per cell - q arm") + 
#   xlab("sum(UMI) - q arm only") +
#   ylab("num cells")
```

What is the distribution of UMI sums for cells?
```{r}
sumTable <- aggregate(UMI~cellIndex, data=chr8matrix, FUN=sum)
colnames(sumTable) <- c("cellIndex", "sumUMI")
ggplot(data = sumTable) + 
  geom_histogram(mapping = aes(x=sumUMI), binwidth = 10) + 
  ggtitle("Sum(UMI) per cell") + 
  xlab("sum(UMI)") +
  ylab("num cells")
```

